Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-16 nips_2018.tex
reorder experiment presentation

\begin{abstract}
  Successful approaches to program induction require a hand-engineered
  domain-specific language (DSL), constraining the space of allowed
  programs and imparting prior knowledge of the domain.  We contribute
  a program induction algorithm called \system that learns a DSL while
  jointly training a neural network to efficiently search for programs
  in the learned DSL.  We use our model to solve symbolic regression
  problems, edit strings, and synthesize functions on lists, showing
  how the model learns a domain-specific library of program components
  for expressing solutions to problems in the domain.
\end{abstract}

@@

\begin{figure*}[h]
  \centering
  \begin{tabular}{rc}\toprule
    \sc{Task}& \begin{tabular}{rcl}
      Nancy FreeHafer&$\longrightarrow$&Dr. Nancy\\
      Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
      Pushmeet Kohli&$\longrightarrow$&???
    \end{tabular}
      \\\midrule
    \sc{Program}& $f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}
      \\\midrule
    \sc{DSL}& \begin{tabular}{l}
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y)))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a)}\\
      \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
      \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}
      \\\bottomrule
  \end{tabular}\right.
\end{figure*}

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we initially provide a generic
 set of programming primitives in a Lisp-like language.
% including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
   &\hspace{0.5cm} \emph{(reverses a list)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
    \vspace{-4em}
 \end{wrapfigure}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to list processing (Section~\ref{textSection}) and text editing (Section~\ref{listSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
and initially provide the system with a generic
set of list processing primitives: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsection{List Functions}\label{listSection}
Synthesizing programs that manipulate data structures is
a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of
learning functions that manipulate lists.
We created 225 Lisp-style list manipulation tasks,
each with 15 input/output examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions:
many of the functions are very complicated,
and the agent must learn to solve these complicated problems from only 225 tasks.
Our data set primarily consists of arithmetic operations upon sequences,
and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.

@@

@@

  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
  \\\bottomrule
  \end{tabular}
\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved
  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}

>>>>

\begin{abstract}
  Successful approaches to program induction require a hand-engineered
  domain-specific language (DSL), constraining the space of allowed
  programs and imparting prior knowledge of the domain.  We contribute
  a program induction algorithm called \system that learns a DSL while
  jointly training a neural network to efficiently search for programs
  in the learned DSL.  We use our model to synthesize functions on lists,
  edit strings, and solve symbolic regression problems, showing how the
  model learns a domain-specific library of program components for
  expressing solutions to problems in the domain.
\end{abstract}

@@

\begin{figure*}[h]
  \centering
  \begin{tabular}{rc}\toprule
    \sc{Task}& \begin{tabular}{rcl}
      Nancy FreeHafer&$\longrightarrow$&Dr. Nancy\\
      Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
      Pushmeet Kohli&$\longrightarrow$&???
    \end{tabular}
      \\\midrule
    \sc{Program}& $f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}
      \\\midrule
    \sc{DSL}& \begin{tabular}{l}
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y)))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a)}\\
      \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
      \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}
      \\\bottomrule
  \end{tabular}
\end{figure*}

@@

We apply \system to three domains:
Lisp-style functions on lists;
FlashFill-style~\cite{gulwani2011automating} string processing; and
symbolic regression;
For each of these we initially provide a generic
set of programming primitives in a Lisp-like language. % TODO: we don't actually give lisp primitives for regression..
% including conditionals, variables, and higher-order functions.
Our algorithm then discovers
its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
   &\hspace{0.5cm} \emph{(reverses a list)}
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
    \vspace{-4em}
 \end{wrapfigure}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to list processing (Section~\ref{listSection}) and text editing (Section~\ref{textSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
and initially provide the system with a generic
set of list processing primitives: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

\subsection{List Functions}\label{listSection}
Synthesizing programs that manipulate data structures is
a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of
learning functions that manipulate lists.
We created 244 Lisp-style list manipulation tasks,
each with 15 input/output examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions:
many of the functions are very complicated,
and the agent must learn to solve these complicated problems from only 244 tasks.
Our data set primarily consists of arithmetic operations upon sequences,
and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}, \code{2},
\code{3}, \code{4}, \code{5}.

@@

  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s

@@

  \\\bottomrule
  \end{tabular}
\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved
  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
